---
title: "Response Rate Estimation, Chile Reentry Study"
output: rmarkdown::github_document
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

In this report, we estimate response rates using the following criteria:

- The estimation by wave only uses cases far beyond the wave-specific observation window. For instance, for the wave *2-months*, we only consider those women who have been in the study for 4 months.
- We use the observed response rates to simulate the final response rate of the study, and the rate 0.70 for the last wave (*one-year*).
- This estimation is based on the administrative records of the study.

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "plots/attrition-"
)
```


```{r, warning = FALSE, include = FALSE}
rm(list=ls(all=TRUE))

path <- "/Users/sdaza/Dropbox/Projects/re-entry/10 investigadores/sdaza/data/records/"

library(readxl)
library(sdazar)
library(stringr)
library(lubridate)
library(ggplot2)
library(ggthemes)

#+ get data
dat <- data.table(read_excel(paste0(path, "20170628 Logros.xlsx"), sheet = "LISTADO TOTAL", skip = 0))
setnames(dat, "ID FOLIO", "id")
anyDuplicated(dat[, id])

text <- lookvar(dat, "NOMBRE")
text <- c(text, lookvar(dat, "APELLIDOS"))
dat[, (text) := NULL]

```

```{r, include = FALSE}
#+ function to clean dates

cleanDates <- function(text) {

  a <- data.table(text)
  a[, temp := str_extract(text, "([0-9]+/[0-9]+/[0-9]+)|([0-9]+-[0-9]+-[0-9]+)")]
  a[, temp := str_replace_all(temp, "-", "/")]
  a[grepl("20[0-9]+", temp), cc := as.Date(temp, format = "%d/%m/%Y")]
  a[is.na(cc) && !grep("20[0-9]+", temp), cc := as.Date(temp, format = "%d/%m/%y")]
  a[is.na(cc) && grep("20[0-9]+", temp), cc := as.Date(temp, format = "%d/%m/%Y")]
  a[is.na(cc) && !grep("20[0-9]+", temp), cc := as.Date(temp, format = "%y/%m/%d")]
  a[is.na(cc) && grepl("20[0-9]+", temp), cc := as.Date(temp, format = "%Y/%m/%d")]
  suppressWarnings(a[as.numeric(text) > 0,
                     cc := as.Date(as.numeric(text), origin = "1899-12-30")])

  a[, year := year(cc)]
  a[, month := month(cc)]
  a[, day := day(cc)]

  a[month %in% c(9:12), year := 2016]
  a[month %in% c(1:8), year := 2017]
  suppressWarnings(a[, nd := ymd(paste(year, month, day, sep= "-"))])

  return(a$nd)
}

```


# Start date

We use the date of release to define the start date in the study. Below, a plot with the number of women who began the study by month.

```{r}
#+ define start date
setnames(dat, "FECHA EGRESO DEFINITIVA", "egreso")
setnames(dat, "FECHA EGRESO", "oegreso")

dat[, start := as.Date(egreso, format = "%Y/%m/%d")] # adjust format of date

# clean dates, see function at the bottom of the document
dat[is.na(start), start := cleanDates(oegreso)]
summary(dat$start)

dat[, month := month(start)]
dat[, year := year(start)]
dat[, date := paste0(year, "-", month)]

# define month-year factor to plot
omonths <- paste0(c(rep(2016, length(9:12)), rep(2017, length(1:7))), "-", c(9:12,1:7))
dat[, date := factor(date, levels = omonths )]

agg <- dat[, .N, date]
ggplot(agg[!is.na(date)], aes(y = N, x = date)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(x = "\nMonth", y = "Number of women\n") +
  theme_minimal()
```

# Dates by wave and rates


## Baseline

7 cases with no dates, 5 inconsistent.

```{r}
#+ baseline
# response
text <- lookvar(dat, "Sí: se realizó Línea base")
setnames(dat, text, "c1")
dat[, c1 := ifelse(grepl("^s", c1, ignore.case =  TRUE), 1, 0)]
table(dat$c1, useNA = "ifany")

# interview date
text <- lookvar(dat, "FECHA ENTREVISTA.+Primera fecha")
setnames(dat, text, "dc1")

dat[, ndc1 := cleanDates(dc1)]
dat[dc1 == "12-01.2017", ndc1 := as.Date("2017-01-12")]
dat[dc1 == "V14 y M18/10", ndc1 := as.Date("2016-10-18")]
dat[dc1 == "V16/12716", ndc1 := as.Date("2016-12-27")]

summary(dat$ndc1)
```

To check these cases:

```{r}
dat[, today := today()]

#  baseline time (in days) with respect to start
t <- as.numeric(dat$ndc1 - dat$start) # difference
Mean(t[t <= 0]) # in days
Min(t[t <= 0]) # is that possible?
table(t > 0) # only few cases with positive numbers
dat[t > 0, .(id, start, ndc1, dc1)]
```

## First Week

8 cases missing, 5 inconsistent.

```{r}
#+ first week
# response
setnames(dat, "PARTICIPA", "c2")
dat[, c2 := ifelse(grepl("^s", c2, ignore.case =  TRUE), 1, 0)]
table(dat$c2, useNA = "ifany")
prop.table(table(dat$c2, useNA = "ifany"))

# interview date
setnames(dat, "FECHA ENTREVISTA", "dc2")
dat[, ndc2 := cleanDates(dc2)]
dat[dc2 == "S29/10 y W02/11", ndc2 := as.Date("2016-11-02")]
summary(dat[c2 ==1, ndc2])
```

```{r}
dat[, week := start +  (7 * 5)] # add 5 weeks to compute rate (not useful, just doing it for completeness)
t <- as.numeric((dat$ndc2 - dat$start)/7) # difference in weeks

Mean(t[t>0])
Max(t[t>0])  # is this right?
Min(t[t>0])

dat[t <= 0, .(id, start, ndc2, dc2)] # check these cases

(tab <- table(dat[today > week | c2 == 1, c2])) # all the sample
prop.table(tab)
rate2 <- prop.table(tab)[2]
```

## Two months

8 cases missing, 1 inconsistent.

```{r}
#+ two months
# response
setnames(dat, "PARTICIPA__1", "c3")
dat[, c3 := ifelse(grepl("^s", c3, ignore.case =  TRUE), 1, 0)]
table(dat$c3, useNA = "ifany")
prop.table(table(dat$c3, useNA = "ifany"))

# interview date
setnames(dat, "FECHA ENTREVISTA__1", "dc3")
dat[, ndc3 := cleanDates(dc3)]
dat[dc3 == "13-0-2017", ndc3 := as.Date("2017-04-13")] # I guess!
dat[dc3 == "27/2", ndc3 := as.Date("2017-02-27")]
summary(dat[c3 == 1, ndc3])
```

```{r}
t <- as.numeric((dat$ndc3 - dat$start) / 30.5) # in months

Mean(t[t>0])
Max(t[t>0])  # 4 months?
Min(t[t>0])

dat[t < 0, .(id, start, ndc3,dc3)]

dat[, twomonths := start + 30.5 * 4] # add 4 months

length(dat[today > twomonths | c3 == 1, c3]) # almost all!
(tab <- table(dat[today > twomonths | c3 == 1, c3]))
prop.table(tab)
rate3 <- prop.table(tab)[2]
```


## Six months

3 missing dates, no inconsistencies!

```{r}
#+ six months
# response
setnames(dat, "PARTICIPA__2", "c4")
dat[, c4 := ifelse(grepl("^s", c4, ignore.case =  TRUE), 1, 0)]
table(dat$c4, useNA = "ifany")
prop.table(table(dat$c4, useNA = "ifany"))

# interview date
setnames(dat, "FECHA ENTREVISTA__2", "dc4")
dat[, ndc4 := cleanDates(dc4)]
dat[ndc4 == "0207-05-15", ndc4 := as.Date("2017-05-15")]
dat[ndc4 == "3017-04-24 ", ndc4 := as.Date("2017-04-24 ")]
summary(dat[c4 == 1, ndc4])
```

```{r}

t <- as.numeric((dat$ndc4 - dat$start) / 30.5)

Mean(t[t>0])
Max(t[t>0])
Min(t[t>0]) # 4 months, really?

dat[t < 0, .(id, start, ndc1, ndc2, ndc3, ndc4)]

dat[, sixmonths := start + 30.5 * 8] # add 8 months

(tab <- table(dat[today > sixmonths, c4]))
prop.table(tab)

rate4 <- prop.table(tab)[2]
```

```{r, warning=FALSE, message=FALSE}
# plot time by application
t2 <- as.numeric((dat$ndc2 - dat$start))
t3 <- as.numeric((dat$ndc3 - dat$start))
t4 <- as.numeric((dat$ndc4 - dat$start))

t <- data.table(id = dat$id, "1 week" = t2, "2 months" = t3, "6  months" = t4)
t <- melt(t, id.vars = "id", value.name = "days", variable.name = "wave")
t[, wave := factor(wave)]
table(t$wave)

ggplot(t, aes(days, fill = wave, colour = wave)) +
  geom_density(alpha = 0.1) + theme_minimal() + theme(legend.position="top") +
   xlim(0, 365) + labs(x = "\nDays from prison release",
   y = "Density\n", title = "Density of time to interview from prison release")+
  geom_vline(xintercept = 7, color = "gray", size = 0.7, linetype = "dotted") +
  geom_vline(xintercept = 30.5 * 2, color = "gray", size = 0.7, linetype = "dotted") +
  geom_vline(xintercept = 30.5 * 6, color = "gray", size = 0.7, linetype = "dotted") +
  annotate("text", x = 7, y = 0.13, label = "1 week", size = 3) +
  annotate("text", x = 30.5 * 2, y = 0.07, label = "2 months", size = 3) +
  annotate("text", x = 30.5 * 6, y = 0.04, label = "6 months", size = 3)
```


## Summary response rates

From first week to six months.

```{r}
# print rates
cbind(rate2, rate3, rate4)
```

## Simulate final response rate

All the effort should be put during these last waves!

```{r}

fcases <- list()
pcases <- list()
fup <- list()

for (i in 1:1000) {

  dat[today < sixmonths & c4 != 1, c4_s := ifelse(runif(.N) < rate4, 1, 0)]
  dat[c4 == 1, c4_s := 1]
  dat[is.na(c4_s), c4_s := c4]

  dat[, c5_s := 0]
  dat[c4_s == 1, c5_s := ifelse(runif(.N) < .70, 1, 0)] # using rate .70
  fcases[[i]] <- table(dat$c5_s)[2]
  pcases[[i]] <- prop.table(table(dat$c5_s))[2] # respect to the total

  agg <- dat[, .N, .(c1,c2,c3,c4_s,c5_s)]
  s <- sum(agg$N)
  agg[, prop := N / sum(N)]
  agg[, followups := apply(.SD, 1, sum), .SDcols = c("c2", "c3", "c4_s","c5_s")]
  fup[[i]] <- agg[followups > 2, sum(N)]/ s

}

```

The expect number of cases in the final sample:

```{r}
summary(unlist(fcases))
```

Cumulative response rate by the final wave:

```{r}
summary(unlist(pcases))
```

Expected proportion of cases with 2 or more waves:

```{r}
summary(unlist(fup))
```

# Appendix

## Function to clean dates

```{r}
cleanDates <- function(text) {

  a <- data.table(text)
  a[, temp := str_extract(text, "([0-9]+/[0-9]+/[0-9]+)|([0-9]+-[0-9]+-[0-9]+)")]
  a[, temp := str_replace_all(temp, "-", "/")]
  a[grepl("20[0-9]+", temp), cc := as.Date(temp, format = "%d/%m/%Y")]
  a[is.na(cc) && !grep("20[0-9]+", temp), cc := as.Date(temp, format = "%d/%m/%y")]
  a[is.na(cc) && grep("20[0-9]+", temp), cc := as.Date(temp, format = "%d/%m/%Y")]
  a[is.na(cc) && !grep("20[0-9]+", temp), cc := as.Date(temp, format = "%y/%m/%d")]
  a[is.na(cc) && grepl("20[0-9]+", temp), cc := as.Date(temp, format = "%Y/%m/%d")]
  suppressWarnings(a[as.numeric(text) > 0,
                     cc := as.Date(as.numeric(text), origin = "1899-12-30")])

  a[, year := year(cc)]
  a[, month := month(cc)]
  a[, day := day(cc)]

  a[month %in% c(9:12), year := 2016]
  a[month %in% c(1:8), year := 2017]
  suppressWarnings(a[, nd := ymd(paste(yeıar, month, day, sep= "-"))])

  return(a$nd)
}
```

```{r, include = FALSE}

save(dat, file = "/Users/sdaza/Dropbox/Projects/re-entry/10 investigadores/sdaza/data/records/register.Rdata")

```
